# Data

## Description

The data sources we have are pulled from multiple sources, including Organization for Economic Cooperation and Development (OECD), United Nations (UN), Statistical Office of the European Union (Eurostat), and World Bank Group. For OECD, UN, and Eurostat, they are responsible for the data collection and the data is collected from administrative data sources, such as civil registration and population and housing censuses. For the World Bank Group, they collect data by compiling from officially-recognized international sources. All of the data are either csv files or excel files. Data from OECD, UN, and Eurostat is updated annually, while the data from the World Bank group is updated quarterly. The dimensions for data collected from Eurostat, UN, OECD, and the World Bank group are 52 rows and 13 columns,  322 rows and 22 columns, 55 rows and 64 columns, and 68 columns and 397,936 rows, respectively. The main issue we have for the data is about organization. Since the data we have collected are pulled from multiple sources, it is rather hard to integrate these data into one file since each data has a different format and contains diverse information. Thus, organizing the data into suitable format for processing is a challenge for us. Lastly, our plan on importing the data is to download the files from the official website first, then transform them into the same format using Excel, and import the transformed files in R for further processing and analysis at last.


Links to our data sources:

https://unstats.un.org/unsd/demographic-social/products/dyb/#statistics

https://www.oecd.org/en/data/datasets/oecd-family-database.html

https://ec.europa.eu/eurostat/web/population-demography/demography-population-stock-balance/database

https://datacatalog.worldbank.org/search/dataset/0037712/World-Development-Indicators 


## Missing value analysis
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
```

```{r}
marriage_data <- read.csv("marriage-rate-per-1000-inhabitants.csv")
divorce_data <- read.csv("divorces-per-1000-people.csv")
# indicator <- read.csv("WDICSV.csv")
```

```{r}
marriage_divorce <- marriage_data |> 
  left_join(divorce_data, by = c("Entity", "Code", "Year")) 
names(marriage_divorce)[1] <- "Country"
names(marriage_divorce)[4] <- "MarriageRate"
names(marriage_divorce)[5] <- "DivorceRate"

na_divorce <- sum(is.na(marriage_divorce$DivorceRate))
na_marriage <- sum(is.na(marriage_divorce$MarriageRate))
```
We still need time to figuring out columns (factors that might influence fertility rates) we need, so here,
we start the missing value analysis by focusing on the Marriage Rate and Divorce Rate data. For non-rich countries the data is sparse, so we currently choosing some specific countries for analyzing. The total sum of missing values in divorce rate data set is `r na_divorce`, the total sum of missing values in marriage rate data set is `r na_marriage`. After knowing the total sum, we can then proceed for detailed analyzation.

2.2 Missing value analysis
```{r}
all_years <- data.frame(Year = 1960:2018)
entities <- unique(marriage_divorce[, c("Country", "Code")])
expanded_data <- merge(entities, all_years, by = NULL)
complete_data <- merge(expanded_data, marriage_divorce, 
                       by = c("Country", "Code", "Year"), 
                       all.x = TRUE)
na_merge_divorce <- sum(is.na(complete_data$DivorceRate))
na_merge_marriage <- sum(is.na(complete_data$MarriageRate))
```
After filling out years that might not included inside the data set that will cause misleading, we have new discovery on missing values. The total sum of missing values in divorce rate column is `r na_merge_divorce`, the total sum of missing values in marriage rate column is `r na_merge_marriage`. 
```{r}
binMat <- rbind(
  Missing = colSums(is.na(complete_data)),
  NonMissing = colSums(!is.na(complete_data))
)

barplot(binMat, 
        main = "Missing Values Current Features", 
        xlab = "Features", 
        ylab = "Count", 
        col = c("#4dffd2", "#ff9999"))

legend("topright", 
       legend = c("Missing values", "Non-Missing values"), 
       fill = c("#4dffd2", "#ff9999"))
```
Firstly, we did a missing value visualization of marriage and divorce data after combining them together as shown in the bar plot.

```{r}
long_data <- complete_data %>%
  pivot_longer(cols = c("MarriageRate", "DivorceRate"), # Include only relevant columns
               names_to = "Variable", values_to = "Value") %>%
  mutate(Missing = is.na(Value)) 

ggplot(long_data, aes(x = Year, y = Country, color = Missing)) +
  geom_point(alpha = 0.6, size = 2) +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "blue"),
                     labels = c("Available", "Missing")) +
  facet_wrap(~Variable, scales = "free_y") +
  labs(title = "Missing Values Across Years and Countries",
       x = "Year", y = "Country", color = "Data Status") +
  theme_minimal() +
  theme(axis.text.x = element_text(),
        strip.text = element_text(face = "bold"))
```
We then delve in detail how the missing value is spreading in different countries and figured 
out that United State, United Kingdom, South Korea, Norway, and Italy have more available data 
that can be used for analyzing. While Singapore, Peru, Bolivia and Argentina seems to have a lot
of missing values in both Marriage Rate and Divorce rate.
