# Data

## Description

The data sources we have are pulled from multiple sources, including Organization for Economic Cooperation and Development (OECD), United Nations (UN), Statistical Office of the European Union (Eurostat), and World Bank Group. For OECD, UN, and Eurostat, they are responsible for the data collection and the data is collected from administrative data sources, such as civil registration and population and housing censuses. For the World Bank Group, they collect data by compiling from officially-recognized international sources. All of the data are either csv files or excel files. Data from OECD, UN, and Eurostat is updated annually, while the data from the World Bank group is updated quarterly. The dimensions for data collected from Eurostat, UN, OECD, and the World Bank group are 52 rows and 13 columns,  322 rows and 22 columns, 55 rows and 64 columns, and 68 columns and 397,936 rows, respectively. The main issue we have for the data is about organization. Since the data we have collected are pulled from multiple sources, it is rather hard to integrate these data into one file since each data has a different format and contains diverse information. Thus, organizing the data into suitable format for processing is a challenge for us. Lastly, our plan on importing the data is to download the files from the official website first, then transform them into the same format using Excel, and import the transformed files in R for further processing and analysis at last.


Links to our data sources:

https://unstats.un.org/unsd/demographic-social/products/dyb/#statistics

https://www.oecd.org/en/data/datasets/oecd-family-database.html

https://ec.europa.eu/eurostat/web/population-demography/demography-population-stock-balance/database

https://datacatalog.worldbank.org/search/dataset/0037712/World-Development-Indicators 


## Missing value analysis
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(naniar)
library(reshape2)
```

```{r}
marriage_data <- read.csv("marriage-rate-per-1000-inhabitants.csv")
divorce_data <- read.csv("divorces-per-1000-people.csv")
education_data <- read.csv("education_long.csv")
fertility_data <- read.csv("fertility_long.csv")
gdp_data <- read.csv("gdp_long.csv")
life_data <- read.csv("life_long.csv")
```

```{r,warning = FALSE}
marriage_divorce <- marriage_data |> 
  left_join(divorce_data, by = c("Entity", "Code", "Year")) 
names(marriage_divorce)[1] <- "Country"
names(marriage_divorce)[4] <- "MarriageRate"
names(marriage_divorce)[5] <- "DivorceRate"

filtered_education <- education_data |> 
  filter(Country %in% marriage_divorce$Country & Year %in% marriage_divorce$Year)

filtered_fertility <- fertility_data |> 
  filter(Country %in% marriage_divorce$Country & Year %in% marriage_divorce$Year)

filtered_gdp <- gdp_data |> 
  filter(Country %in% marriage_divorce$Country & Year %in% marriage_divorce$Year)

filtered_life <- life_data |> 
  filter(Country %in% marriage_divorce$Country & Year %in% marriage_divorce$Year)

final_data <- marriage_divorce |>
  left_join(filtered_education, by = c("Country", "Code", "Year")) |>
  left_join(filtered_fertility, by = c("Country", "Code", "Year")) |>
  left_join(filtered_gdp, by = c("Country", "Code", "Year")) |>
  left_join(filtered_life, by = c("Country", "Code", "Year"))

colSums(is.na(final_data))
```
The above result shows how many missing value are there in each feature. After knowing the exact value, we can then doing visualization for better analyzing missing patterns.

```{r}
missing_values <- colSums(is.na(final_data))

missing_data_df <- data.frame(
  Column = names(missing_values),
  MissingCount = missing_values
)

missing_proportion <- missing_values / nrow(final_data)
missing_data_df$Proportion <- missing_proportion

ggplot(missing_data_df, aes(x = reorder(Column, -Proportion), y = Proportion)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Proportion of Missing Values per Column",
       x = "Columns",
       y = "Proportion of Missing Values") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
We merged all relevant features with fertility rate into our final data set and first doing a bar plot visualizing porprotion of missing value in each column. Educational_attainment have the highest proportion of missing values, exceeding 50%. Columns like DivorceRate have a moderate proportion of missing values (around 20–40%). Columns such as gdp_growth, fertility_rate, Gender, life_expectancy, Code, Country, MarriageRate, and Year appear to have minimal or no missing values, making them more reliable for immediate use. 
```{r}
time_missing <- final_data %>%
  group_by(Year) %>%
  summarise(MissingCount = sum(is.na(.))) 

ggplot(time_missing, aes(x = Year, y = MissingCount)) +
  geom_line(color = "red") +
  geom_point() +
  labs(title = "Missing Values Over Time", x = "Year", y = "Count of Missing Values") +
  theme_minimal()
```
We also see if the number of missing value changes with time. From the plot above, the count of missing values appears nearly constant across all years in the data set. There are no visible spikes or drops in missing values over time.This consistency suggests that missingness is not influenced by temporal factors. The uniformity could indicate systematic missingness, where data is consistently missing for specific features, countries, or regions across all years.